Learning to Pivot with Adversarial Networks

### 2017-05-17 tex/nips/nips_2017.tex
save

Assuming a probability model $p(X,Y,Z)$, where $X$ are the data, $Y$ are the
target labels, and $Z$ are the nuisance parameters, we consider the problem of
learning a predictive model $f(X)$ for $Y$ conditional on the observed values of $X$
that is robust to uncertainty in the unknown value of $Z$. We
introduce a flexible learning procedure based on adversarial networks~\citep{goodfellow2014generative}
for enforcing that  $f(X)$ is a pivot with respect to $Z$.
We derive theoretical results proving that the procedure
converges towards a model that is both optimal and statistically independent of the
nuisance parameters (if that model exists) or for which one can tune a
trade-off between accuracy and robustness (e.g., as driven by a higher level objective).
In particular, and to the best of our knowledge, our contribution is the first
working solution for imposing statistical independence on the predictive model, regardless of the type of
the nuisance parameter (discrete or continuous) or of its prior.
Finally, we demonstrate the
effectiveness of the approach with a toy example and an example from particle physics.

@@

While our contribution may first appear as a minor extension, most previous
works are based on the strong and limiting assumption that $Z$ is a binary
random variable (e.g., $Z=0$ for the source domain, and
$Z=1$ for the discrete domain). For
instance, the motivation behind \citep{pan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,zemel2013learning,ganin2014unsupervised,ajakan2014domain,EdwardsS15,louizos2015variational}
takes root in the minimization of some form of divergence between the two distributions of
$f(X)|Z=0$ and $f(X)|Z=1$. For this reason, and in contrast with our procedure,
these works cannot intrinsically be generalized to non-binary or
continuous nuisance parameters, both from a practical and theoretical point of view.
Of notable interest however, \cite{kamishima2012fairness} enforces fairness through
a prejudice regularization term based on empirical estimates of $p(f(X)|Z)$.
While this approach is in principle sufficient for handling non-binary
nuisance parameters $Z$, it requires accurate
empirical estimates of $p(f(X)|Z=z)$ for all values $z$, which quickly becomes
impractical as the cardinality of $Z$ increases. By contrast, our approach models the conditional
dependence through an adversarial network, which allows for generalization without
necessarily requiring a growing number of training examples.

>>>>

Assuming a probability model $p(X,Y,Z)$, where $X$ are the data, $Y$ are the
target labels, and $Z$ are the nuisance parameters, we consider the problem of
learning a predictive model $f(X)$ for $Y$ conditional on the observed values of $X$
that is robust to uncertainty in the unknown value of $Z$. We
introduce a flexible learning procedure based on adversarial networks~\citep{goodfellow2014generative}
for enforcing that  $f(X)$ is a pivot with respect to $Z$.
We derive theoretical results proving that the procedure
converges towards a model that is both optimal and statistically independent of the
nuisance parameters (if that model exists) or for which one can tune a
trade-off between accuracy and robustness (e.g., as driven by a higher level objective).
In particular, and to the best of our knowledge, our contribution is the first
solution for imposing pivotal constraints on a predictive model,
working regardless of the type of
the nuisance parameter (discrete or continuous) or of its prior.
Finally, we demonstrate the
effectiveness of the approach with a toy example and an example from particle physics.

@@

While our contribution may first appear as a minor extension, most previous
works are based on the strong and limiting assumption that $Z$ is a binary
random variable (e.g., $Z=0$ for the source domain, and
$Z=1$ for the target domain). In particular, \citep{pan2011domain,gong2013connecting,baktashmotlagh2013unsupervised,zemel2013learning,ganin2014unsupervised,ajakan2014domain,EdwardsS15,louizos2015variational}
are all based on the minimization of some form of divergence between the two distributions of
$f(X)|Z=0$ and $f(X)|Z=1$. For this reason, and in contrast with our procedure,
these works cannot intrinsically be generalized to non-binary or
continuous nuisance parameters, both from a practical and theoretical point of view.
Of notable interest however, \cite{kamishima2012fairness} enforces fairness through
a prejudice regularization term based on empirical estimates of $p(f(X)|Z)$.
While this approach is in principle sufficient for handling non-binary
nuisance parameters $Z$, it requires accurate
empirical estimates of $p(f(X)|Z=z)$ for all values $z$, which quickly becomes
impractical as the cardinality of $Z$ increases. By contrast, our approach models the conditional
dependence through an adversarial network, which allows for generalization without
necessarily requiring a growing number of training examples.
